---
title: "Extra "
output: html_notebook
---
# Adding interaction terms 


```{r}
#With interactions
bp_full_int <- lm(data=bp_train, sbp ~ gender + married + smoke + exercise + age + weight + height + race + alcohol + trt  + stress + salt + chldbear + income + educatn + overwt + income:educatn + gender:chldbear + weight:gender + height:gender)

summary(bp_full_int)
```

- since $R^2$ is really low for blood pressure, we need to add more terms. 
- added interaction terms to see if $R^2$ goes up significantly or not.


## Residual diagnostics of the interaction model

```{r}
qqnorm(bp_full_int$residuals)
qqline(bp_full_int$residuals)
hist(bp_full_int$residuals)
#augment(bp_full)
ggplot(bp_full_int, aes(x=.fitted, y=.resid)) + geom_point() + geom_hline(yintercept = 0)
boxcox(bp_full_int)
shapiro.test(bp_full_int$residuals)
bptest(bp_full_int)
```
- This model looks like it has residuals that are


## AIC/BIC with interaction terms
```{r}
stepAIC(bp_full_int, direction = "backward")
#AIC=1934.13
```



```{r}
bp_reduced_int <- lm(sbp ~ gender + smoke + exercise + age + weight + alcohol + trt + 
    stress + salt + chldbear + income + educatn + income:educatn + 
    weight:stress + age:income + smoke:alcohol + exercise:salt, data=bp_train)
summary(bp_reduced_int)
```

## PRESS Statistic for interaction model

```{r}
PRESS(bp_reduced3)
```




## Multicollinearity
VIF:
```{r}
R_int <- summary(bp_full_int)$r.squared
VIF_int <- 1/(1-R_int)
VIF_int
```
- low VIF, not likely to have a multicollinearity problem with this model

```{r}
anova(bp_reduced_int, bp_full_int)
```
- With interaction terms model -- P-value > 0.05, reduced model is sufficient


## Cross Validation for interaction model
```{r}
# fit model for training set
# gender + married + smoke + exercise + age + height + race + 
#    alcohol + trt + bmi + stress + chldbear + income + educatn + 
#    overwt + income:educatn + trt:overwt + married:stress + 
#    smoke:exercise + alcohol:educatn

bp_valid %>% mutate(`income:educatn` = income*educatn, `trt:overwt` = trt*overwt, `married:stress` = married*stress, `smoke:exercise` = smoke*exercise, `alcohol:educatn` = alcohol*educatn) -> bp_valid_int
anova(bp_reduced_int)

pred_cols_int <- c('gender', 'married', 'smoke', 'exercise', 'age', 'height', 'race', 'alcohol' , 'trt', 'bmi', 'stress', 'chldbear', 'income', 'educatn', 'overwt', 'income:educatn', 'trt:overwt', 'smoke:exercise', 'alcohol:educatn')

pred_bp_int <- predict(bp_reduced_int, bp_valid_int[pred_cols_int])
delta_bp_int <- bp_valid_int['sbp'] - pred_bp_int
n.star_int <- dim(bp_valid_int)[1]
MSPR_int <- sum((delta_bp_int)^2)/n.star_int 
MSPR_int
```

- MSE = 607.6
- MSPR = 666.486


```{r}
PRESS(lm_AIC_int)
```


## Graphical Diagnostics 

```{r}
p2_int <- ols_plot_cooksd_chart(bp_reduced)
p3_int <- ols_plot_dffits(bp_reduced)
p4_int <- ols_plot_resid_lev(bp_reduced)
p5_int <- ols_plot_resid_stud_fit(bp_reduced)
p6_int <- ols_plot_dfbetas(bp_reduced)
```
- We have a couple of outliers, however majority of these values are within
the threshold 
- observation 263 gives us the highest influence, it is an outlier


## Remedial measures


