---
title: "BloodPressure"
output: html_notebook
---

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(MASS)
library(onewaytests)
library(readxl)
```

# Reading and cleaning the model
```{r}
bp <- read_xlsx("BloodPressure.xlsx")
bp %>% mutate(married = replace(married, married == 'N', 0))%>% 
  mutate(married = replace(married, married == 'Y', 1)) %>%
  mutate(gender = replace(gender, gender == 'F', 0)) %>%
  mutate(gender = replace(gender, gender == 'M', 1)) %>%
  mutate(smoke = replace(smoke, smoke == 'N', 0)) %>%
  mutate(smoke = replace(smoke, smoke =='Y', 1))-> bp

class(bp$gender) = "double"
class(bp$married) = "double"
class(bp$smoke) = "double"
bp
```
```{r}
bp_full <- lm(data=bp, sbp ~ gender + married + smoke + exercise + age + weight + height + race + alcohol + trt + bmi + stress + salt + chldbear + income + educatn)
summary(bp_full)
```

# Residual diagnostics
```{r}
qqnorm(bp_full$residuals)
qqline(bp_full$residuals)
hist(bp_full$residuals)
augment(bp_full)
ggplot(bp_full, aes(x=.fitted, y=.resid)) + geom_point() + geom_hline(yintercept = 0)
boxcox(bp_full)
shapiro.test(bp_full$residuals)
# add bf test here; didnt work earlier
```

- p-value > 0.05, we can conclude the error looks from a normal population
- fitted vs residual values look randomly scattered
- normal qq plot looks like its in a stright line
- data seems to be cleaned up here.


# Graphs of SBP and covariates
```{r}
bp  %>% pivot_longer(c(gender:educatn), names_to = "xnames", values_to = "x")%>% 
  ggplot(aes(x=x, y=(sbp)))  + geom_point() + facet_wrap(~xnames, scales = "free") + geom_smooth()
```


# Checks for multicollinearity
```{r}
# add graphs for correlation matrix
bp %>%  dplyr::select(-sbp) -> bp_cor
cor(bp_cor)
```


# Backwards elimination

Which criterion is better?


# AIC/BIC
# $C_p$ 
# $SSE_p$ / $R^2$
# $MSE_p$ / $R^2_{adj}$
# PRESS